---
title: "R Notebook"
output: html_notebook
---


6E1
State three motivating criteria that define information entropy. Try to express each in your own words.
```{r}

```

6E2
Suppose a coin is weighted such that, when it is tossed and lands on a table, it comes up with heads 70% of the time. What is the entropy of this coin?
```{r}
p <- c(0.3, 0.7)
entropy <- -sum(p*log(p))
entropy
```
6E3
Suppose that a four sided die i loaded such that, when tossed onto a table, it shows "1" 20% of the time, "2" 25% of the time, "3" 25% of the time, "4" 30% of the time. What is the entropy of this die?



```{r}
p <- c(0.2, 0.25, 0.25, 0.3)
entropy <- -sum(p*log(p))
entropy
```

6E4
```{r}
p <- c(1/3, 1/3, 1/3)
entropy <- -sum(p*log(p))
entropy
```

6M1
AIC:

DIC:

WAIC:

6M2
Explain the difference between model selection and model averaging. What information is lost under model selection? What information is lost under model averaging?

6M3

When comparing models with an information criterion, why must all models be fit to exactly the same observations? What would happen to the information criterion values, if the models were fit to different numbers of observations? Perform some experiments if you are not sure.

6M4
What happens to the effective number of parameters, as measured by DIC or WAIC, as a prior becomes more concentrated? Why? Perform some experiments, if you are not sure.

6M5 Provide an informal explanation of why informative priors reduce overfitting.

6M6
Provide and information explanation of why overly informative priors result in overfitting.



6H1
```{r}
library(rethinking)
data(Howell1)
d <- Howell1
d$age <- (d$age - mean(d$age))/ sd(d$age)
d$age2 <- d$age ** 2
d$age3 <- d$age ** 3
d$age4 <- d$age ** 4
d$age5 <- d$age ** 5
d$age6 <- d$age ** 6
set.seed(1000)
i <- sample(1:nrow(d), size = nrow(d) / 2)
d1 <- d[i,]
d2 <- d[-i,]

```

```{r}
m6H1.1 <- map(
  alist (
    height ~ dnorm(mu, sigma),
    mu <- a + b1*age ,
    a ~ dnorm(0,100),
    b1 ~ dnorm(0,100),
    sigma ~ dunif(0,50)
  ),
  data = d1
)
```

```{r}
m6H1.2 <- map(
  alist (
    height ~ dnorm(mu, sigma),
    mu <- a + b1*age + b2*age^2 ,
    a ~ dnorm(0,100),
    b1 ~ dnorm(0,100),
    b2 ~ dnorm(0,100),
    sigma ~ dunif(0,50)
  ),
  data = d1
)
```

```{r}
m6H1.3 <- map(
  alist (
    height ~ dnorm(mu, sigma),
    mu <- a + b1*age + b2*age^2 + b3*age^3 ,
    a ~ dnorm(0,100),
    b1 ~ dnorm(0,100),
    b2 ~ dnorm(0,100),
    b3 ~ dnorm(0,100),
    sigma ~ dunif(0,50)
  ),
  data = d1
)
```

```{r}
m6H1.4 <- map(
  alist (
    height ~ dnorm(mu, sigma),
    mu <- a + b1*age + b2*age^2 + b3*age^3 + b4*age^4,
    a ~ dnorm(0,100),
    b1 ~ dnorm(0,100),
    b2 ~ dnorm(0,100),
    b3 ~ dnorm(0,100),
    b4 ~ dnorm(0,100),
    sigma ~ dunif(0,50)
  ),
  data = d1
)
```

```{r}
m6H1.5 <- map(
  alist (
    height ~ dnorm(mu, sigma),
    mu <- a + b1*age + b2*age^2 + b3*age^3 + b4*age^4 + b5*age^5,
    a ~ dnorm(0,100),
    b1 ~ dnorm(0,100),
    b2 ~ dnorm(0,100),
    b3 ~ dnorm(0,100),
    b4 ~ dnorm(0,100),
    b5 ~ dnorm(0,100),
    sigma ~ dunif(0,50)
  ),
  data = d1
)
```

```{r}
m6H1.6 <- map(
  alist (
    height ~ dnorm(mu, sigma),
    mu <- a + b1*age + b2*age2 + b3*age3 + b4*age4 + b5*age5 + b6*age6,
    a ~ dnorm(0, 100),
    b1 ~ dnorm(0,100),
    b2 ~ dnorm(0,100),
    b3 ~ dnorm(0,100),
    b4 ~ dnorm(0,100),
    b5 ~ dnorm(0,100),
    b6 ~ dnorm(0,100),
    sigma ~ dunif(0,50)
  ),
  data = d1
)
```

```{r}
plot(precis(m6H1.1))
```

```{r}
plot(precis(m6H1.6))
```
```{r}
WAIC(m6H1.1)
```

```{r}
WAIC(m6H1.2)
```

```{r}
(models <- compare(m6H1.1,m6H1.2,m6H1.3,m6H1.4,m6H1.5,m6H1.6))
```

```{r}
plot(models)
```

