---
title: "Statistical Rethinking"
output: html_notebook
author: Magnus Gasslander
email: magnus.gasslander@gmail.com
---

Excercises from the Statistical Rethinking book

```{r}
library(rethinking)
```

## 2E1
$$Pr(rain|Monday)$$


## 2E2

The probability that it is Monday, given that it is raining.

## 2E3

$$Pr(Monday|rain)$$

and

$$Pr(rain|Monday)*Pr(Monday) / Pr(rain)$$

## 2E4

It is just our small world that deals with propabilities. In the real world there are no probabilities only truth

## 2M1

1 (W,W,W)

```{r}
p_grid <- seq(from = 0, to = 1, length.out = 20)
prior <- rep(1,20)
likelihood <- dbinom(x = 3, size = 3, prob = p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior)
plot(p_grid,posterior)
```

2 (W,W,W,L)

```{r}
p_grid <- seq(from = 0, to = 1, length.out = 20)
prior <- rep(1,20)
likelihood <- dbinom(x = 3, size = 4, prob = p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior)
plot(p_grid,posterior)
```
3 (L,W,W,L,W,W,W)

```{r}
p_grid <- seq(from = 0, to = 1, length.out = 20)
prior <- rep(1,20)
likelihood <- dbinom(x = 5, size = 7, prob = p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior)
plot(p_grid,posterior)
```
## 2M2

```{r}
p_grid <- seq(from = 0, to = 1, length.out = 20)
prior <- ifelse(p_grid < 0.5,0,1)
likelihood <- dbinom(x = 5, size = 7, prob = p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior)
plot(p_grid,posterior)
```

## 2M3

Show that $Pr(Earth|Land) = .23$

$$Pr(Earth|Land) = \frac {Pr(Land|Earth)*Pr(Earth)} {Pr(Land)}$$
```{r}
.3 * .5 /(.5*1 + .5*.3)
```
.3 * .5 / 0.65 = 0.23

## 2M4

BB
BB
BW
WB
WW
WW

There are three ways to choose a B front card. 2 of those has a black back side

p(BB|B) = 2/3

## 2M5

BB
BB
BB
BB
BW
WB
WW
WW

There are 5 ways to choose a black front. 4 of those have a black back side

P(BB|B) = 4/5

## 2M6
BB
BB
(below twice as common as BB)
BW
BW
WB
WB
(below 3 times as common as BB)
WW
WW
WW
WW
WW
WW

There are 4 ways to pull a B front side card. Of those there are 2 ways to get a black back side



## 2M7

BB -> -BW +WB +WW +WW

BB -> -BW +WB +WW +WW

BW -> -BB -BB +WW +WW

WB -> -BB -BB +WW +WW

WW -> -BB -BB -BW +WB

WW -> -BB -BB -BW +WB


There are total 8 options to pull a B front then a white front.

6 of those have the BB card

p(B front, W front) = 6/8 = 3/4 = .75



## 2H1

$$p(twins, twins) = p(A|twins)*p(twins|A) + p(B|twins)*p(twins|B)$$

$$p(twins, twins) = (1/3)*.1 + (2/3)*.2$$


```{r}
p_twins_twins <- (1/3)*.1 + (2/3)*.2
p_twins_twins
```

## 2H2

$$p(A|twins) = \frac {p(twins/A) * p(A)}  {p(twins)}$$
```{r}
p_A_twins <- .1*.5 / (.1*.5 + .2*.5)
p_A_twins
```


## 2H3

calculate $P(twins, single)$

$$p(A|twins, single) = \frac {p(twins, single|A)* p(A)}  {p(twins, single)}$$

$$p(A|twins, single) = .1*.9*.5 /  .1*.9*.5 + .2*.8*.5$$

```{r}
p_A_twins_single <- .1*.9*.5 / (.1*.9*.5 + .2*.8*.5)
p_A_twins_single
```

Alternative method using p(A|twins) as prior

$$p(A|twins, single) =\frac{ p(single|A) * p(A|twins)} {p(single)}$$

prior = p(At) = p(A|twins)

$$p(A|twins, single) = \frac {p(single|A)*p(At)} {p(single|A)*p(At) + p(single|B) * p(Bt)}$$

$$p(A|twins, single) = .9 * (1/3)/(.9*(1/3) + .8*(2/3))$$

```{r}
p_A_twins_single <- .9*(1/3)/(.9*(1/3) + .8*(2/3))
p_A_twins_single
```
## 2H4

$$p(Atest|A) = .8$$
$$p(Btest|B) = .65$$

$$p(A|Atest) = \frac {p(Atest|A)*p(A)} {p(Atest|A)*p(A) + p(Atest|B)*p(B)}$$
```{r}
p_A_Atest <- .8*.5/(.8*.5 + .35*.5)
p_A_Atest
```

### 2H4(2)

$$\frac {p(Atest|A)*p(A)} {p(Atest|A)*p(A) + p(Atest|B)*p(B)}$$

```{r}
#Hm, check
p_A_Atest_update <- .8*.36/(.8*.36 + .35*.64)
p_A_Atest_update
```


# Chapter 3


```{r}
p_grid <- seq(from = 0, to = 1, length.out = 1000)
prior <- rep (1, 1000)
likelihood <- dbinom (6, size = 9, prob = p_grid)
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample(p_grid, prob = posterior, size = 1e4, replace = TRUE)
hist(samples)

```

## 3E1
```{r}
sum(samples < .2) / 1e4
```

## 3E2
```{r}
sum(samples > .8) / 1e4
```
## 3E3
```{r}
sum(samples < .8 & samples > .2) / 1e4
```
## 3E4
```{r}
quantile(samples, .2)
```
## 3E5
```{r}
d <- quantile(samples, .8)
d
```

## 3E6
```{r}
HPDI(samples, .66)
```

## 3E7
```{r}
quantile(samples , c(.17, .83))
```

```{r}
PI(samples , .66)
```
## 3M1
```{r}
p_grid <- seq(from = 0, to = 1, length.out = 20)
prior <- rep(1,20)
likelihood <- dbinom(x = 8, size = 15, prob = p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior)
plot(p_grid,posterior)
```

## 3M2
```{r}
samples <- sample(p_grid, prob = posterior, size = 1e4, replace = TRUE)
HPDI(samples, .9)
```

## 3M3
```{r}
w <- rbinom(10000, size=15, prob = samples)
simplehist(w)
prob_8 <- sum(w == 8) / 1e4
prob_8
```

## 3M4
```{r}
w9 <- rbinom(10000, size=9, prob = samples)
prob_6 <- sum(w9 == 6) / 10000
prob_6
```

## 3M5
```{r}
p_grid <- seq(from = 0, to = 1, length.out = 20)
prior <- c(rep(0,10), rep(1,10))
likelihood <- dbinom(x = 8, size = 15, prob = p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior)
plot(p_grid,posterior)
```
```{r}
samples <- sample(p_grid, prob = posterior, size = 1e4, replace = TRUE)
HPDI(samples, .9)
```


```{r}
w <- rbinom(10000, size=15, prob = samples)
simplehist(w)
prob_8 <- sum(w == 8) / 1e4
prob_8
```
```{r}
w9 <- rbinom(10000, size=9, prob = samples)
prob_6 <- sum(w9 == 6) / 10000
prob_6
```


## 3H1


```{r}
library(rethinking)
data(homeworkch3)
sum(birth1) + sum(birth2)

p_grid <- seq(from = 0, to = 1, length.out = 100)
prior <- rep(1, 100)
likelihood <- dbinom (111, size = 200, prob = p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior)
plot(p_grid, posterior, type = "b" )

p_grid[which.max(posterior)]


```
## 3H2
```{r}
samples <- sample(p_grid, size = 10000, replace = TRUE, prob = posterior)
HPDI(samples, prob= .5)
HPDI(samples, prob = .89)
HPDI(samples, prob = .97)
```


## 3H3
```{r}
s <- rbinom(10000, 200, samples)
#sampls <- data.frame(xx = c(s,samples), yy = rep(letters[1:2],each = 10000))
#ggplot(sampls, aes(x=xx, fill=yy)) + geom_histogram(position="identity", alpha=0.2)
dens(s, show.HPDI=.89)
abline(v=sum(c(birth1, birth2)))
```
The constructed posterior distribution is wider 

## 3H4
```{r}
s <- rbinom(10000, 100, samples)
#hist(s)
dens(s, show.HPDI=.89)
abline(v=sum(birth1))

```
## 3H5
Count number of first borns who were girls
```{r}
first_born_girls <- sum(birth1 == 0)
p_grid <- seq(from = 0, to = 1, length.out = 100)
prior <- rep(1, 100)
likelihood <- dbinom (first_born_girls, size = 100, prob = p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior)
plot(p_grid, posterior, type = "b" )


```

```{r}
samples <- sample(p_grid, size = 10000, replace = TRUE, prob = posterior)
s <- rbinom(10000, 100, samples)
dens(s, show.HPDI=.89)
abline(v=sum(birth2))
```

If they get a boy the first time they stop getting more children and then they will not be part of this group (hm, does this logic work?)